Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-03-25 GWLDCII.tex
split into two papers

\title{Combinatorics of the geometry of Wilson loop diagrams}
\author{Susama Agarwala, Si\^an Fryer, and Karen Yeats}
%\date{}

@@

This paper studies the combinatorics of Wilson Loop Diagrams.
\section{Wilson Loop diagrams}\label{section background}

What are Wilson loop diagrams and their integrals.

@@

There is one particular type subdiagram that deserves special attention.

\begin{dfn}
For $W$ an admissible diagram, $(P, V(P))$ is exact if $|V(P)| = |P| + 3$.
\end{dfn}

The exact subdiagrams define an equivalence relation amongst Wilson loop diagrams.

\begin{dfn}\label{equivdfn} 
There is an equivalence relationship on the set of admissible Wilson loops diagrams given by the transititve closure the following binary relation: $W = (\cP, [n]) \sim W'= (\cP', n)$ if
\begin{enumerate}
\item There exist two different exact subdiagrams, $(P, V(P))$ and $(P', V(P'))$ of $W$ and $W'$ respectively such that $V(P) =  V(P')$.
\item The complementary subdiagrams are identical: $(\cP \setminus P, V(P)^c) = (\cP' \setminus P', V(P')^c)$.
\end{enumerate}
\end{dfn}

\begin{eg} \label{eg:equivdiags}
Note that since this is an equivalence relation, we may find that two Wilson loop diagrams are equivalent, even if they do not have complements of (non-trivial) exact subdiagrams in common. Consider the following three Wilson loop diagrams,
\bas
W_1 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{2}{0}{4}{-1}
    \drawprop{5}{0}{8}{0}
    \drawprop{5}{1}{7}{0}
		\end{scope}
	\end{tikzpicture} \quad
W_2 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{2}{0}{4}{-1}
    \drawprop{5}{0}{8}{0}
    \drawprop{6}{0}{8}{-1}
		\end{scope}
	\end{tikzpicture} \quad
W_3 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{1}{1}{3}{0}
    \drawprop{5}{0}{8}{0}
    \drawprop{6}{0}{8}{-1}
		\end{scope}
	\end{tikzpicture} .
\eas

The diagrams $W_1 \sim W_2$ because $(\{(5,8), (5,7)\}, \{5,6,7,8,1\})$ and $(\{(5,8), (7,8)\}, \{5,6,7,8,1\})$ are the corresponding differing subdiagrams. Furthermore, there is the equivalence $W_2 \sim W_3$ due to the exact subdiagrams $(\{(1,4), (3,4)\}, \{1,2,3,4,5\})$ and $(\{(1,4), (1,3)\}, \{1,2,3,4,5\})$. This forces an equivalence between $W_1$ and $W_3$, even though one cannot partition the propagators of each into a an exact subdiagram (that may vary between the diagrams) and a complement that is fixed.  \end{eg}

@@

\section{Equivalence classes of Wilson loop diagrams}

In \cite{wilsonloop}, Agarwala and Amat show that Wilson loop diagrams can be interpreted as positroids, a certain well behaved class of realizable matroids (this correspondence is stated precisely in Theorem \ref{thm WLD defines matroid} below). This opens up the study of Wilson loop diagrams to techniques from geometry and combinatorics. In particular we examine the partial result shown in \cite{wilsonloop}, that shows that equivalent Wilson Loop Diagrams define the same matroid. In section 2.4, we make the relationship between Wilson Loop diagrams and matroids explicit. We count the size of each equivalence class, and enumerate the {\color{violet} the number of said classes}. We also show that two Wilson Loop diagrams map to the same matroid if and only if they are equivalent.

In subsection \ref{sec matroid background}, we discuss a few matroidal facts about Wilson loop diagrams. In section \ref{sec: polygon partitions}, we show that there is a one to one correspondence between exact subdiagrams of Wilson loop diagrams and triangulated pieces of the corresponding polygon partion. In section \ref{sec: exact diagram matroidal props}, we prove some matroidal properties of exact subdiagrams. We show that a subdiagram of $W$ defines a uniform matroid if and only if the subdiagram is exact (Theorem \ref{exactuniformthm}). Furthermore, this uniform matroid is the restriction of the dual matriod of $W$ restricted to the exact subdiagram (Remark \ref{remark exact dual restiction}). The main result of section \ref{sec: matroids and equivalence} shows that that two admissible Wilson loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). We obtain a formula for the number of admissible Wilson loop diagrams in each equivalence class (Corollary \ref{number of equiv diagrams}) {\color{violet}as well as the number of equivalence classes}. In this way, we give a comple characterization of the relationship between Wilson Loop diagrams and their associated matroids.

\subsection{Wilson loop diagrams as matroids\label{sec matroid background}}

We first give a quick summary of the matroid terminology that we will need; it is not intended as a comprehensive introduction to matroids and the interested reader is referred to [ref][find a good matroid reference].

A {\em matroid} $M = (E,\cB)$ consists of a finite ground set $E$ and a non-empty family $\cB \subseteq \cP(E)$ whose elements satisfy the {\em basis exchange property}: for any distinct $B_1,B_2 \in \cB$ and any $a \in B_1 \setminus B_2$, there exists some $b \in B_2 \setminus B_1$ such that $(B_1 \setminus \{a\})\cup \{b\} \in \cB$ as well. The elements of $\cB$ are called the {\em bases} of the matroid. Note that the basis exchange property immediately implies that all bases have the same size.

A subset $A \subseteq E$ is called {\em independent} in $M$ if $A \subseteq B$ for some $B \in \cB$, and {\em dependent} else. The {\em rank}  $\rk(A)$ of a subset $A \subseteq E$ is the size of the largest independent set contained in $A$. The rank of the matroid itself is defined to be $\rk(E)$.

A {\em circuit} in $M$ is a minimally dependent set. That is, it is a set $C \subseteq E$ such that $C$ is dependent but $C \setminus \{e\}$ is independent for any $e \in C$. A union of circuits is called a {\em cycle}. On the other hand, a {\em flat} is a maximally dependent set, i.e. a set $F \subseteq E$ such that $\rk(F \cup \{e\}) = \rk(F) + 1$ for any $e \in E \setminus F$. Unsurprisingly, a {\em cyclic flat} is a set which is both a flat and a cycle. The set of circuits in a matroid uniquely defines that matroid, as does the set of flats; thus one could specify a matroid by listing its independent sets, bases, circuits, or flats. [ref]

Finally, we describe several important types of matroids. A matroid of rank $k$ with a ground set of size $n$ is called {\em realizable} if there exists some $A \in Gr(k,n)$ whose non-zero $k\times k$ minors are exactly those with columns indexed by elements of $\cB$. A {\em positroid} is a matroid which can be realized by an element of the totally nonnegative Grassmannian $\Gr(k,n)$. Finally, a {\em uniform matroid} of rank $r$ is a matroid in which any set of size $\leq r$ is independent.

Matroid theory relates to the study of Wilson loop diagrams as follows. In \cite{wilsonloop}, Agarwala and Amat show that every admissible Wilson loop diagram with $k$ propagators defines a positroid of rank $k$, and that the independent sets can be read directly from the diagram:

\begin{thm} \label{thm WLD defines matroid} \cite[Theorem 3.6]{wilsonloop} Any admissible Wilson loop diagram $W =(\cP, [n])$ defines a matroid $M(W)$ with ground set $[n]$. The independent sets are exactly those subsets $V \subseteq [n]$ such that $\nexists U\subseteq V$ satisfying $|\Prop(U)| < |U|$. \label{thm:WLDmatroid}\end{thm}
In other words, the independent sets of $M(W)$ correspond to the sets of vertices in $W$ such that no subset supports fewer propagators than the vertices it contains.

Throughout, we take the {\em matroid defined by $W$} to be the matroid $M(W)$ of Theorem \ref{thm WLD defines matroid}. Note that since vertices of the diagram $W$ correspond to columns of the associated matrix $C(W)$, $M(W)$ can also be thought of as the matroid realized by $C(W)$.

% The identification of Theorem \ref{thm WLD defines matroid} is not 1-1: by \cite[Theorem 1.18]{wilsonloop}, if two admissible Wilson loop diagrams are equivalent (as in Definition~\ref{equivdfn}) then they define the same matroid.

Let $W = (\cP,n)$ be an admissible Wilson loop diagram, and $M(W)$ its associated matroid. Where it will not cause confusion we conflate the two objects, identifying vertices of $W$ with elements of the ground set $[n]$ in $M(W)$.

In particular, this allows us to prove results about $M(W)$ by considering the behavior of propagators in $W$. We record a few elementary facts about the rank and cycles of $M(W)$ here as an example of this.

\begin{lem}\label{lem facts about WLD matroids}
Let $W = (\cP,n)$ be an admissible Wilson loop diagram. Then:
\begin{enumerate}
\item The rank of a set $V \subseteq [n]$ is bounded above by $\min\{|V|,|\Prop(V)|\}$, with $\rk(V) = |V|$ if and only if $V$ is an independent set.
%\item Let $v \in [n]$ and $q \in \Prop(v)$. Then for any $V \subset [n]$ such that $q \not \in \Prop(V)$, we have $\rk(V \cup \{v\}) = \rk(V) +1$, i.e. adding a vertex that supports a new propagator to a set increases the rank of the set.
\item If $C \subseteq [n]$ is a cycle, then $\rk(C) = |Prop(C)|$.
\item If $[n]$ can be partitioned into at least two non-empty sets, each of which support different sets of propagators that form a partition of the propagotor set, \bas [n] = \sqcup_i V(P_i) \quad \textrm{s.t.} \quad \sqcup P_i = \cP; \quad V(P_i) \cap V(P_j) = \emptyset; \quad  P_i \cap P_j = \emptyset \;, \eas then the matroid $M(W)$ is seperable, \bas M(W) = \bigoplus_i M(P_i, V(P_i)) \;.\eas 
\item $F(P)$ is a flat of $M(W)$, thus justifying the name ``propagator flat''.
\end{enumerate}
\end{lem}
\begin{proof}
The first part of (1) is \cite[Equation (9)]{wilsonloop} and surrounding discussion, and the second part is standard matroid theory. (2) is \cite[Lemma 3.27]{wilsonloop}. (3) is a direct consequence of \cite[Lemma 3.20]{wilsonloop} and the fact that $F(P_1)^c = V(P_1^c)$

To prove (4), we need to show that $F(P)$ is maximally dependent. If $F(P) = [n]$ then this is automatic, so suppose not and let $v \in [n] \setminus F(P)$. In other words, $v \in V(P^c)$ and so $v$ supports some propagator $q \not\in P$.  Let $S \subseteq F(P)$ be an independent set of maximal size. Then $\Prop(S) \subseteq P$ by (1), and no subset of $S$ supports fewer propagators than the number of vertices it contains (this is the definition of an independent set in $M(W)$). Since $v$ supports a new propagator $q \not\in P$, the set $S \cup \{v\} \subseteq F(P) \cup\{v\}$ also satisfies this independence condition. Thus $\rk(F(P) \cup\{v\}) = \rk(F(P)) + 1$, as required.
\end{proof}

Note that this means that the set, $F(\emptyset)$, of non-supporting vertices is the maximal subset of vertices of $W$ of rank $0$. That is, it is the unique flat of rank $0$ in $M(W)$.

\subsection{Polygon partitions of Wilson loop diagrams\label{sec: polygon partitions}}

The equivalence relation on Wilson loop diagrams is defined in terms of exact subdiagrams; thus in order to understand the equivalence, we need a way to extract and compare exact subdiagrams. We do this via the notion of a polygon partition of $W$.

\begin{dfn}
  Let $W = (\cP, [n])$ be an admissible Wilson loop diagram.  The \emph{polygon partition} associated to $W$, denoted $\tau(W)$, is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a chord edge of $\tau(W)$; specifically,  a propagator $(i,j) \in \cP$ defines a chord connecting the vertices $i$ and $j$ in $\tau(W)$.
  \end{itemize}
\end{dfn}

\begin{lem}\label{tausimpleplanarlem}
If $W = (\cP, [n])$ is an admissible Wilson loop diagram, then $\tau(W)$ is a simple planar graph whose outer face is a cycle. It is embedded such that the vertices all lie on this infinite face\footnote{That is, it is an \emph{outerplanar} graph}. These vertices are cyclically ordered, with a choice of first vertex giving it an additional compatible linear order.
\end{lem}

\begin{proof}
Since the vertices of $\tau(W)$ are labeled by the edges of $W$, which are cyclically ordered, this gives an ordering to the vertices and the outer face of $\tau(W)$ is a cycle. Since $W$ is admissible, no pairs of propagators cross. Therefore, it is a planar embedding. Similarly, $W$ does not admit any propagators of the form $p = (i, i+1)$; therefore there is exactly one edge connecting any two adjacent edges of $\tau(W)$. Finally, there does not exist two propagators $p,q$ such that both $p$ and $q$ start at edge $i$ and end at edge $j$. Therefore, no other two vertices of $\tau(W)$ can be connected by more than one edge. Finally, the embedding of $\tau(W)$ is induced from the embedding of the graph $W$.
 \end{proof}

\begin{comment}
\hlfix{For an admissible $W$ no propagators cross and so no chord edges of $\tau(W)$ cross.  Thus, in graph theoretic language, we can view $\tau(W)$ as a planar embedding of a
graph with no cut vertices, where all vertices lie on the infinite face, along with a distinguished start vertex and direction around the infinite face.  From this viewpoint the polygon of $tau(W)$ is the boundary of the infinite face.}{I've smushed this together a bit in the prev. lem. Tell me if I've missed anything. Its a lemma for emphasis, not difficulty}
\end{comment}

\begin{eg}\label{WLDtopolygonpartition}
In this example we return to two of the Wilson loop diagrams in Example \ref{eg:equivdiags}. We can pair diagrams with their polygon partitions as follows:
\bas W_1 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{2}{0}{4}{-1}
    \drawprop{5}{0}{8}{0}
    \drawprop{5}{1}{7}{0}
		\end{scope}
	\end{tikzpicture} \quad; \quad
\tau(W_1) = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{8}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{2}{4}
    \drawchord{5}{8}
    \drawchord{5}{7}
	\end{scope}
	\end{tikzpicture}
\eas and
\bas W_3 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
    \drawprop{1}{0}{4}{0}
	\drawprop{1}{1}{3}{0}
    \drawprop{5}{0}{8}{0}
    \drawprop{6}{0}{8}{-1}	
		\end{scope}
	\end{tikzpicture}\quad; \quad
\tau(W_3) = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{8}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{1}{3}
    \drawchord{5}{8}
    \drawchord{6}{8}
	\end{scope}
	\end{tikzpicture} .
\eas

Recall that a planar embedding of a graph is a \emph{triangulation} if all faces, except possibly the infinite face, are triangles.

\begin{dfn}
  Let $W$ be an admissible Wilson loop diagram and $\tau(W)$ its polygon partition. A \emph{triangulated piece} of $\tau(W)$ is a 2-connected subgraph of $\tau(W)$ which is a triangulation. We will take the convention that a subgraph consisting of a single chord edge is called a \emph{trivial} triangulated piece.
A {\em maximal} triangulated piece is one which is not contained in any strictly larger triangulated piece.
\end{dfn}

\begin{dfn}
 A {\em decomposition} of a polygon partition $\tau(W)$ is a set of 2-connected induced subgraphs of $\tau(W)$ which partition the edges of $\tau(W)$.  
\end{dfn}

\begin{eg} \label{eg: unique decomposition} For the Wilson loop diagrams and polygon partitions in Example \ref{WLDtopolygonpartition}, the vertex sets $\{1, 2, 3,4\}$ and $\{5, 6, 7, 8\}$ give maximal triangulated pieces for both $\tau(W_1)$ and $\tau(W_3)$. The vertex set $\{4,5, 8, 1\}$ is not a triangulation in either polygon partition. 
\end{eg}

\begin{lem} \label{decompositionlem}
  For $W$ an admissible Wilson loop diagram, the polygon partition $\tau(W)$ has a unique decomposition into maximal triangulated pieces, and edges in the polygon of $\tau(W)$.
\end{lem}

\begin{proof}
We begin by giving an algorithm for the decomposition, then prove its uniqueness. Let $W = (\cP, [n])$, with $|\cP| = k$.

By \emph{splitting} a vertex $v$ we will mean replacing $v$ by new vertice $v_1, v_2,\ldots, v_{\text{deg}(v)}$ such that each $v_i$ has exactly one neighbour and the union of the of the $v_i$ is the neighbourhood\footnote{The neighbourhood of a vertex is the set of adjacent vertices} of $v$.

Let $T(W)$ be the dual graph of $\tau(W)$ with the vertex corresponding to the infinite face split.
%In other words, place a vertex on each finite face of $\tau(W)$. These vertices are connected if there is a chord edge separating the edges. Furthermore, if the face is bounded by an edge of the polygonal cycle of $\tau(W)$, the corresponding vertex gets a leaf edge for each such boundary.
Since $\tau(W)$ is an embedded graph (with a fixed distinguished embedding) by Lemma \ref{tausimpleplanarlem}, $T(W)$ is an uniquely defined graph.

Furthermore $T(W)$ is a tree because it is connected, has $n+k+1$ vertices ($k+1$ from the internal faces of $\tau(W)$ and $n$ from the outer face) and $n+k$ edges (since $\tau(W)$ has $n+k$ edges).  Additionally, since $\tau(W)$ is simple, $T(W)$ has no vertices of degree $2$.

%We claim that $T(W)$ is a tree. Since $\tau(W)$ is a planar graph with $k$ chord edges and no internal vertices, there are $k+1$ internal faces of $\tau(W)$. Each of these internal faces corresponds to a vertex of $T(W)$, and there are $n$ vertices of $T(W)$ from the splitting of the dual graph, so $T(W)$ has $n+k+1$ vertices in total. On the other hand, $T(W)$ has $n+k$ edges, corresponding to the $n +k$ edges of $\tau(W)$. Therefore, $T(W)$ is a tree.

Split every vertex of $T(W)$ which has degree $>3$.
%By construction, each face of $\tau(W)$ is a polygon. Therefore, each vertex of $T(W)$ is at least 3 valent (one for each edge of the polygon). \hlfix{If split each vertex of $T(W)$ if it is strictly greater than $3$ valent.}{???} (In other words, we fail to split exactly when the face is a triangle).
The connected components of $T(W)$ correspond to the decomposition of $\tau(W)$ into maximal triangulated pieces and edges originally in the polygon of $\tau(W)$.  Let $f$ be the forest thus obtained.
The vertices of $f$ either have degree 1 or 3.
%, (if they correspond leaf vertices) or 3,  (if they correspond to triangular faces).
Trees of $f$ with no trivalent vertices correspond to either edges in the polygon of $\tau(W)$, if they were originally leaves of $T(W)$, or to maximal trivial triangulated pieces.
Splitting at all the faces that are not triangles ensures maximality of the decomposition. If the splitting were not maximal, then one could add a triangle to a connected component of the splitting, but this would imply that that splitting happened at a valence $3$ vertex.

\bas 
\begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{8}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{1}{3}
    \drawchord{5}{8}
    \drawchord{6}{8}
 \draw (0,0) node {$\bullet$};
\draw (-1,.2) node {$\bullet$};	
\draw (-.5,1.1) node {$\bullet$};
\draw (.8,-.3) node {$\bullet$};	
\draw (1.1,-.5) node {$\bullet$};
\draw(1.1,-.5) -- (.8,-.3) --(0,0)--(-1,.2) -- (-.5,1.1);
\foreach \i in {1,2,...,8} {
      \draw (45*\i:2) node {$\bullet$};
    }
\draw (45*1:2) --(0,0) -- (45*5:2);
\draw (45*6:2) -- (.8,-.3);
\draw (45*8:2) -- (1.1,-.5) -- (45*7:2);
\draw (45*2:2) -- (-.5,1.1) -- (45*3:2);
\draw (45*4:2) -- (-1,.2);
	\end{scope}
	\end{tikzpicture} \rightarrow 
\begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{8}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{1}{3}
    \drawchord{5}{8}
    \drawchord{6}{8}
 \draw (.1,.1) node {$\bullet$};
\draw (.1,-.1) node {$\bullet$};
\draw (-.1,-.1) node {$\bullet$};
\draw (-.1,.1) node {$\bullet$};
\draw (-1,.2) node {$\bullet$};	
\draw (-.5,1.1) node {$\bullet$};
\draw (.8,-.3) node {$\bullet$};	
\draw (1.1,-.5) node {$\bullet$};
\draw(1.1,-.5) -- (.8,-.3) --(.1,-.1); 
\draw (-.1, .1) --(-1,.2) -- (-.5,1.1);
\foreach \i in {1,2,...,8} {
      \draw (45*\i:2) node {$\bullet$};
    }
\draw (45*1:2) --(.1,.1);
\draw (-.1, -.1) -- (45*5:2);
\draw (45*6:2) -- (.8,-.3);
\draw (45*8:2) -- (1.1,-.5) -- (45*7:2);
\draw (45*2:2) -- (-.5,1.1) -- (45*3:2);
\draw (45*4:2) -- (-1,.2);
	\end{scope}
	\end{tikzpicture}
\eas

To see uniqueness, consider a different maximal decomposition of $\tau(W)$. This induces a splitting on $T(W)$, where each connected component of the new decomposition corresponds to a subtree. Call this forest $f'$. Since $f' \neq f$, there are two trees, $t$ and $t'$ in $f$ and $f'$ that are distinct, but share at least one edge of $T(W)$. Since $f'$ is also maximal, $t'$ is not a subtree of $t$. Therefore, the edges of $t'$ can be found in at least two trees in the forest $f$. In particular, there is a vertex $v$ in $t'$ that corresponds to a split vertex of $T(W)$ in the original decomposition.
This implies that $v$ has valence greater than $3$ in $T(W)$, and thus the corresponding face of $\tau(W)$ is not a triangle. In other words, the decomposition corresponding to $f'$ is not a triangulation. 
\end{proof}

\begin{cor} \label{maxtriangdisjointcor}
Given a maximal decomposition of $\tau(W)$, the maximal triangulated pieces are edge disjoint.
\end{cor}

\begin{proof}
Consider any two distinct maximal triangulated pieces of $\tau(W)$. These two pieces correspond to subtrees of $T(W)$ and intersect, at most, at a vertex in the interior of $\tau(W)$. Since the subtrees corresponding to the maximal triangulated pieces are edge disjoint, and the edges of $T(W)$ correspond to the edges of $\tau(W)$, this forces the maximal triangulated pieces to be edge disjoint as well.
\end{proof}

We are now in a position to relate the triangulated pieces of $\tau(W)$ to exact subdiagrams of $W$.

Every triangulated piece $t$ of an admissible Wilson loop diagram $W$ corresponds to a subdiagram of $W$ by taking the set of propagators $P$ corresponding to edges of $t$ and then taking the subdiagram $W_P = (P, V(P))$.  Conversely, given a subdiagram $W_P = (P, V(P))$ of $W$ we can obtain a subgraph of $\tau(W)$, called $t$, as follows:
\begin{itemize}
\item The vertex set of the subgraph $t$ is
  \begin{itemize}
  \item the vertices of $\tau(W)$ corresponding to edges of $W$ defined by cyclically consecutive elements of $V(P)$ as a subset of $[n]$
  \end{itemize}
\item The edge set of the subgraph is
  \begin{itemize}
  \item the edges of $\tau(W)$ corresponding to propagators of $P$
  \item along with the outer edges of $\tau(W)$ for which both their end points are in the vertex set
  \end{itemize}
\end{itemize}
Note that the subgraph $t$ depends on how the propagators of $P$ sit inside $[n]$, not on how they sit in $W_P$ itself.  In particular $t$ is not the subgraph of $\tau(W)$ consisting only of edges corresponding to propagators of $P$.

\begin{lem}\label{lem triang to exact}
  Let $W$ be an admissible Wilson loop diagram and $\tau(W)$ its polygon partition.  The triangulated pieces of $\tau(W)$ correspond to the exact subdiagrams of $W$ using the correspondence described above.
\end{lem}

\begin{proof}
Let us first record a few standard facts about polygon triangulations (that is, about triangulations with all vertices on the outer face).  If such a triangulation has $n$ vertices then it has $n$ edges on the polygon (that is, on the outer face) and $n-3$ edges which are not.  No planar graph with the same vertices and the same outer face can have more edges than the triangulation, and every such simple graph with $n-3$ edges off the outer face is a triangulation.

Since $W$ is admissible, by Lemma \ref{tausimpleplanarlem} $\tau(W)$ is a simple graph. Let $t$ be a triangulated piece of the decomposition of $\tau(W)$ given in Lemma \ref{decompositionlem}; note that $t$ cannot be equal to $\tau(W)$ by the definition of admissible diagrams.

If $t$ has 2 vertices then $t$ corresponds to a propagator that connects two non-adjacent edges. Therefore, the trivial triangulation is a trivial exact subdiagram.

Now suppose that $t$ has $m>2$ vertices.  We count how many edges of $t$ are not on the outer face of $\tau(W)$.  These are exactly the edges of $t$ defined by propagators of $W$. Consider the intersection of $t$ with the outer face of $\tau(W)$: this is a possibly disconnected subgraph of the polygon of $\tau(W)$ and this subgraph has $m$ vertices. Call this new subgraph $S$, and let $j$ be the number of connected components of $S$.   To join the components of $S$ into the outer face of $t$, $t$ must have $j$ edges in its outer face which are not in the outer face of $\tau(W)$.  Furthermore $t$ has $m-3$ edges not in its outer face and so also not in the outer face of $\tau(W)$.  Thus there are $m-3+j$ edges of $t$ not in the outer face of $\tau(W)$.

Each of these $m-3+j$ internal edges corresponds to a propagator in $W$; call this set of propagators $P$.  Next we count the size of $V(P)$.  Each of the $m$ vertices in the outer face of $t$ corresponds to an edge of $W$. These $m$ edges define $j$ connected components of the outer polygon of $W$. Thus the set $V(P)$ has $m+j$ vertices.  In other words, 
\[|V(P)| = m+j = |P| +3\;.\]
Thus the subdiagram $(P,V(P))$ defined by $t$ is exact.

Conversely, suppose we have an exact subdiagram $(P, V(P))$ of $W = (\cP, [n])$ supported on $|V(P)| = |P|+3$ vertices, and let $t$ be the subgraph of $\tau(W)$ corresponding to $(P,V(P))$.

Suppose $|P|=1$. Let $p$ be the element of $P$. The exactness condition on $(P, V(P))$ says that the four supporting vertices of $p$ are distinct.  If the support of $p$ is four consecutive vertices, then $V(p)$ defines three consecutive boundary edges of $W$, so $t$ is a single triangle, hence a triangulated piece.  If the support of $p$ is not four consecutive vertices, then the vertices which are the ends of $t$ are separated by at least two vertices along the cycle.  This implies that $t$ is a trivial triangulated piece.

Now suppose $|P|>1$.  Let $j=|P|$, $m=|V(P)|$, and 
suppose that the set $V(P)$ defines $c$ disjoint cyclic intervals of $[n]$. Then $t$ has $m-c$ vertices.  If $t$ were a triangulation, $t$ would have $j -c$ internal edges.

The graph $t$ has $j$ edges that come from propagators, and $m - 2c$ edges that come from the boundary polygon of $\tau(W)$. Since $t$ has $m-c$ vertices, it has $m-c$ external edges, of which $c$ come from propagators. Therefore, of the $j$ edges of $t$ that come from propagators, $j-c$ are internal to the connected component. Therefore, $t$ is a triangulated piece.

%  The statement of the lemma then follows form the fact that inclusion is preserved under $\tau$ and so maximality also corresponds under $\tau$.
\end{proof}

To avoid the issue of exact diagrams being subdiagrams of other exact subdiagrams (for instance, any subdiagram $(q, V(q))$, for $q \in \cP$ is exact), we introduce the notion of maximal exact subdiagrams.

\begin{dfn}
An exact subdiagram $(P, V(P))$ is a {\em maximal exact subdiagram} of $W$ if there is no other exact subdiagram $(Q, V(Q))$ in $W$ that contains $(P,V(P))$ as a strict subdiagram.
\end{dfn}

\begin{cor} \label{uniqueproppartitioncor}
Any admissible Wilson loop diagram $W = (\cP, [n])$ can be uniquely decomposed into maximal exact subdiagrams. These maximal subdiagrams partition $\cP$.
\end{cor}

\begin{proof}
Combining Lemmas \ref{decompositionlem} and \ref{lem triang to exact} yields the unique decomposition into maximal exact subdiagrams, and Corollary \ref{maxtriangdisjointcor} ensures that no propagator appears in more than one subdiagram in this decomposition. Since the chord edges of $\tau(W)$ correspond to the propagators of $W$, the decomposition of $\tau(W)$ induces a partition of $\cP$.
\end{proof}

\subsection{Matroidal properties of exact subdiagrams \label{sec: exact diagram matroidal props}}

Since Corollary \ref{uniqueproppartitioncor} allows us to decompose any admissible Wilson loop diagram into a collection of maximal exact subdiagrams, in this section we examine the matroid properties of exact subdiagrams more closely. In this section, we show prove two results:
\begin{enumerate}
\item The matriod assoicated to an exact subdiagram of $W$ can be written as a contraction of the matroid $M(W)$ by the complementary propagator flat (Theorem \ref{exact diagrams contractions}). 
\item The matroid associated to an exact subdiagram is uniform (Theorem \ref{exactuniformthm} ).
\end{enumerate}

We begin by proving some useful facts about propagator flats, and flats of matroids associated to admissible Wilson loop diagrams more generally.

\begin{comment}

\begin{dfn}\label{def prop flat} 
Let $W = (\cP,[n])$ be an admissible Wilson loop diagram and $P \subseteq \cP$. The set $F(P) := V(P^c)^c$ is called the {\em propagator flat} of $P$.
\end{dfn}
Thus in Lemma \ref{contractsubdiaglem} above we would contract $M(W)$ by the propagator flat $F(P^c)$ of the {\em complement} of $P$ in order to study $(P,V(P))$. The justification for this notation is given by (1) of the next lemma: the set $F(P)$ consists of vertices which {\em only} support propagators in $P$ (or no propagators at all).

\begin{lem}\label{lem properties of prop flats} \sanote{do we need the first 2? They follow directly from the definition 1.2. Do we use them later?}
Let $F(P)$ be a propagator flat as defined above. Then
\begin{enumerate}
\item $F(\emptyset)$ is exactly the set of vertices which support no propagators, while for $P \neq \emptyset$ we have \[F(P) = \big(V(P) \setminus V(P^c)\big) \cup F(\emptyset).\] \note{updated (1) to match new definition of $F(P)$; }
\item If $Q\subseteq P$ then $F(Q) \subseteq F(P)$.
\item $F(P)$ is a flat of $M(W)$, thus justifying the name ``propagator flat''.
\end{enumerate}
\end{lem} \sanote{I'd like to attach this to the "generic facts about matroids" in lemma 2.2?}
\begin{proof}
The proof of (1) and (2) are routine applications of the definition and are omitted.

>>>>

\title{Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators}
\author{Susama Agarwala, Si\^an Fryer, and Karen Yeats}
%\date{}

@@

\section{Introduction}

***Some general intro to WLDs.  It can probably gloss over most of what is already in the other paper, but this time we are going to talk about the denominator, so for context we should probably say something about the integral, and its poles in this intro.  Possibly also more about the integral should go at the end of the background section.***

In \cite{wilsonloop}, Agarwala and Amat show that Wilson loop diagrams can be interpreted as positroids, a certain well behaved class of realizable matroids.  This opens up the study of Wilson loop diagrams to techniques from geometry and combinatorics. 
In the companion paper to this one ***cite it*** we looked at equivalence classes of diagrams using matroidal and polytopal tools.  However, the combinatorial richness of positroids comes from more than just their matroidal nature, but from their many bijections with other interesting combinatorial structures, including plabic graphs, decorated permutations, and Le diagrams, see for instance \cite{Oh2}.

In this paper we will focus on implications of the Le diagrams on the mathematics and physics of Wilson loop diagrams.

\subsection{Roadmap}

Section~\ref{section general background} gives the required background for Wilson loop diagrams (subsection~\ref{section WLD background}) and for positroids and some of their many combinatorial versions (subsection~\ref{sec:positroid background}).

Section~\ref{sec GN algorithm} proceeds to give an algorithm to go directly from a Wilson loop diagram to its Grassmann necklace (Algorithm~\ref{alg:put GN on WLD} and the proof that it is correct in Theorem~\ref{res:alg gives GN}).  In view of the known combinatorial equivalences, it was already possible to get the Grassmann necklace of a Wilson loop diagram, but it involved going through the matrix $C(W)$ and the positroid as a matroid.  The direct algorithm shows us directly how each propagator contributes to each element of the Grassmann necklace.  This idea is crucial for all the arguments of the present paper.  Along with the algorithm, Section~\ref{sec GN algorithm} contains two key techincal lemmas, Lemma~\ref{lem sian}, the focus of subsection~\ref{sec:propagator configs}, and the cyclic interval lemma, Lemma~\ref{vertex cyclic int lem} in subsection~\ref{sec:GN alg}, which leads into the proof of Theorem~\ref{res:alg gives GN} as well as being important in subsequent sections.

Section~\ref{sec dim} proceeds to show that the dimension of the positroid cell defined by a Wilson loop diagram has dimension three times the number of propagators, Theorem~\ref{thm dim}.  The proof here is very explicit, keeping track of how a propagator contributes $+$s to the Le diagram, and correspondingly somewhat technical.  Section~\ref{sec poles} characterizes the denominator of the integral of a Wilson loop diagram in terms of how propagators contribute to the Grassmann necklace.  This result is also given by an algorithm, Algorithm~\ref{alg WLD to denom via GN}, and theorem, Theorem~\ref{thm denom}.

\section{Background}\label{section general background}

\subsection{Wilson Loop diagrams}\label{section WLD background}

\note{We should be a bit careful about text overlap between the two papers in the WLD background section, as some bits are copied (and potentially also for the intro depending on what we do)}

@@

@@

\subsection{Positroid background}\label{sec:positroid background}
